#Including required packages
```{r}
# 1. Library (must be installed prior to loading
library(RMySQL)     ### MySQL
library(tidyr)      ### tidyr Seperating rows for handling multivalued attributes
library(sqldf)      ### SQLdf for using SQL commands on dataframes
options(sqldf.driver = "SQLite")
library(dplyr)      ### dplyr for splitting dataframes
```

#Reading CSV file
```{r}

#2. CSV File Name:
#setwd("/Users/sanyamharne/Desktop/Practicum1/")
file_csv <- 'BirdStrikesData.csv'

df1 <- read.csv(file_csv)
df2 <- separate_rows(df1, Conditions..Precipitation)

```

#Setting up the database connection
```{r}

# 4. Settings
db_user <- 'root'
db_password <- 'Sanyam@123'
db_name <- 'practicum1'

# Local: localhost
db_host <- 'localhost' 
db_port <- 3306 # always this port unless you change it during installation

# 4. Connect to DB
mydb <-  dbConnect(MySQL(), user = db_user, password = db_password,
                dbname = db_name, host = db_host, port = db_port)

#To send Local Variables we write the below statement
dbSendQuery(mydb, "SET GLOBAL local_infile = true;")

#dbListTables(mydb)
```

#Loading data into aircraft table
```{r}
aircraft <- sqldf('SELECT DISTINCT "Aircraft..Make.Model" AS Make_Model, "Aircraft..Type" AS Type, "Is.Aircraft.Large." AS IsAircraftLarge  FROM df1')
aircraft[aircraft$IsAircraftLarge == "Yes",]$IsAircraftLarge <- 1
aircraft[aircraft$IsAircraftLarge == "No",]$IsAircraftLarge <- 0
#aircraft
#aircraft$AircraftTypeID<- 1:354
aircraft$IsAircraftLarge<- as.integer(aircraft$IsAircraftLarge)
dbWriteTable(mydb, name = "aircraft", value= aircraft, append = TRUE,row.names=FALSE, overwrite = FALSE)
```

#Loading into airpline_operator table
```{r}
airline_operator <- sqldf('SELECT DISTINCT "Aircraft..Airline.Operator" AS AirlineOperator FROM df1')
#airline_operator
#airline_operator$AirlineOperatorID <- 1:293
dbWriteTable(mydb, name = "airline_operator", value= airline_operator, append = TRUE,row.names=FALSE)

```

#Loading into airline_aircrafts table
```{r}
aircraft <- dbGetQuery(mydb, "SELECT * FROM `aircraft`;")

airline_operator <- dbGetQuery(mydb, "SELECT * FROM `airline_operator`;")

airline_aircrafts <- sqldf("SELECT DISTINCT aircraft.AircraftTypeID, airline_operator.airlineOperatorID, df1.`Aircraft..Number.of.engines.` AS `NumberOfEngines` 
                           FROM aircraft INNER JOIN airline_operator INNER JOIN df1
                           WHERE aircraft.`Make_Model` = df1.`Aircraft..Make.Model` AND 
                           airline_operator.AirlineOperator = df1.`Aircraft..Airline.Operator`;")

dbWriteTable(mydb, name = "airline_aircrafts", value= airline_aircrafts, append = TRUE,row.names=FALSE, overwrite = FALSE)
airline_aircrafts
```

#Loading into wildlife table
```{r}
wildlife <- sqldf('SELECT DISTINCT "Wildlife..Species" AS Species, "Wildlife..Size" AS Size 
                  FROM df1')
#If dbWriteTable
#wildlife$WildlifeID <- 1:395
#wildlife
dbWriteTable(mydb, name = "wildlife", value= wildlife, append = TRUE,row.names=FALSE)

```

#Loading into airport table
```{r}
airport <- sqldf('SELECT DISTINCT "Airport..Name" AS NAME, "Origin.State" AS State FROM df1')
#airport
#airport$AirportID<- 1:1141
dbWriteTable(mydb, name = "airport", value= airport, append = TRUE,row.names=FALSE)

```

#Loading data into precipitation_type table
```{r}
precipitation_type <- sqldf('SELECT DISTINCT "Conditions..Precipitation" AS Precipitation_Type 
                            FROM df1 
                            WHERE Precipitation_Type NOT LIKE "%,%"')
#precipitation_type
dbWriteTable(mydb, name = "precipitation_type", value= precipitation_type, append = TRUE,row.names=FALSE)
```


#Loading data into precipitation
```{r}

precipitation_type <- dbGetQuery(mydb, "SELECT * FROM `precipitation_type`;")

precipitation <- sqldf("SELECT df2.`Record.ID`, PrecipitationID 
                       FROM df2 INNER JOIN precipitation_type 
                       ON df2.'Conditions..Precipitation'= precipitation_type.Precipitation_Type;")
dbWriteTable(mydb, name = "precipitation", value= precipitation, append = TRUE,row.names=FALSE)
```




#Loading data into bird_strike_incident table 
```{r}
cname <- c("Record.ID", "FlightDate", "Wildlife..Number.struck", "Wildlife..Number.Struck.Actual", "Effect..Impact.to.flight", "Altitude.bin", "Effect..Indicated.Damage", "When..Phase.of.flight", "Remains.of.wildlife.collected.", "Remains.of.wildlife.sent.to.Smithsonian", "Remarks", "Conditions..Sky", "Pilot.warned.of.birds.or.wildlife.", "Cost..Total..", "Feet.above.ground", "Number.of.people.injured")

bird_strike_incident <- df1 %>% select(all_of(cname))
bird_strike_incident$FlightDate <- as.Date(bird_strike_incident$FlightDate, format = "%m/%d/%Y")

airport <- dbGetQuery(mydb, "SELECT * FROM `airport`;")
bird_strike_incident$AirportID <- 0

x<-sqldf("SELECT DISTINCT `Record.ID`, AirportID 
         FROM df1 INNER JOIN airport 
         ON df1.'Airport..Name' = airport.Name AND df1.`Origin.State` = airport.State ")
x
bird_strike_incident[bird_strike_incident$Record.ID == x$Record.ID, ]$AirportID<-x[bird_strike_incident$Record.ID == x$Record.ID,]$AirportID

wildlife <- dbGetQuery(mydb, "SELECT * FROM `wildlife`;")

bird_strike_incident$WildlifeID<- 0
x<-sqldf("SELECT DISTINCT df1.'Record.ID',wildlife.WildlifeID FROM df1 INNER JOIN wildlife ON df1.`Wildlife..Species`= wildlife.Species AND df1.`Wildlife..Size` = wildlife.Size")
bird_strike_incident[bird_strike_incident$Record.ID == x$Record.ID,]$WildlifeID <- x[bird_strike_incident$Record.ID == x$Record.ID,]$WildlifeID

aircraft <- dbGetQuery(mydb, "SELECT * FROM `aircraft`;")
airline_operator <- dbGetQuery(mydb, "SELECT * FROM `airline_operator`;")
airline_aircrafts <- dbGetQuery(mydb, "SELECT * FROM `airline_aircrafts`;")

bird_strike_incident$PlaneID<- 0
x1 <- sqldf("SELECT DISTINCT PlaneID, Make_Model, AirlineOperator, NumberOfEngines 
            FROM aircraft INNER JOIN airline_operator INNER JOIN airline_aircrafts
            WHERE airline_aircrafts.AircraftTypeID = aircraft.AircraftTypeID AND airline_aircrafts.AirlineOperatorID = airline_operator.AirlineOperatorID")

x<-sqldf("SELECT DISTINCT df1.'Record.ID', x1.`PlaneID`
         FROM df1 INNER JOIN x1
         WHERE df1.`Aircraft..Make.Model` = x1.Make_Model AND df1.`Aircraft..Airline.Operator` = x1.AirlineOperator AND df1.`Aircraft..Number.of.engines.` = x1.NumberOfEngines;")

bird_strike_incident[bird_strike_incident$Record.ID == x$Record.ID,]$PlaneID <- x[bird_strike_incident$Record.ID == x$Record.ID,]$PlaneID
bird_strike_incident
colnames(bird_strike_incident)
colnames(bird_strike_incident) <- c("RecordID", "FlightDate", "WildLife_NumberOfStruck", "Wildlife_NumberOfStruckActual", "Effect_ImpactToFlight", "Altitude_Bin", "EffectIndicatedDamage", "PhaseOfFlight", "WildLifeRemainsCollected", "WildLifeRemainsSent", "Remarks", "ConditionsSky", "PilotWarned", "Cost_Total", "FeetAboveGround", "PeopleInjured", "PlaneID", "WildlifeID", "AirportID")
dbWriteTable(mydb,  name = "bird_strike_incident", value= bird_strike_incident, append = FALSE, row.names=FALSE, overwrite = TRUE)



```



```{r}
print(colnames(df1))
```

```{sql connection=mydb}
SELECT COUNT(*) FROM bird_strike_incident
```
```{sql connection=mydb}
SELECT * FROM aircraft
```
```{sql connection=mydb}
SELECT * FROM airport

```
```{r}
bird_strike_incident
```
```{sql connection=mydb}
SELECT * FROM wildlife

```



```{sql connection=mydb}
SELECT * FROM precipitation
```






Question 4 :  Create a SQL query against your database to find the number of bird strike incidents for each airport upon take-off or climb. Include all airlines. You may either use a {sql} code chunk or an R function to execute the query.
```{sql connection=mydb}
SELECT COUNT(RecordID), A.Name
FROM bird_strike_incident AS B INNER JOIN airport AS A
ON B.AirportID = A.AirportID

GROUP BY B.AirportID

```

```{r}
sqlcmd = "SELECT COUNT(RecordID),A.Name FROM bird_strike_incident AS B INNER JOIN airport AS A ON B.AirportID = A.AirportID
          GROUP BY B.AirportID "
dd<-sqldf(sqlcmd)
dd
```


```{sql connection=mydb}
SELECT COUNT(*) FROM 

```


Question  5 : Create a SQL query against your database to find the airports that had the most bird strike incidents (during any flight phase). Include all airlines. You may either use a {sql} code chunk or an R function to execute the query.
```{sql connection=mydb}
SELECT A.Name , COUNT(B.RecordID)
FROM bird_strike_incident as B INNER JOIN airport as A
ON B.AirportID = A.AirportID
GROUP BY B.AirportID
HAVING COUNT(RecordID) = MAX(COUNT(RecordID))

```

```{sql connection=mydb}
SELECT COUNT(RecordID)
FROM bird_strike_incident
GROUP BY AirportID
```


```{r}
dbDisconnect(mydb)
```







SS